---
title: "[eb1] Analyses"
author: "Egor Ananyev"
date: "2019-01-04"
output:
    html_document:
        toc: true
editor_options: 
  chunk_output_type: console
---

<style type="text/css">
body, td{
  font-family: Cambria;
}
code.r{
  font-family: Inconsolata;
}
</style>

# Data and function input

```{r set-options, message=F}
# options(width=100)
library(plyr)
library(dplyr)
library(ggplot2)
library(lme4)
computer = 'shocky'  # 'shocky', 'station3', 'acer' are all possible options
if(computer == 'shocky' || computer == 'toshi'){ dropbox_dir='/Users/Egor/Dropbox/' }
if(computer == 'acer'){ dropbox_dir='/Users/egora/Dropbox/' }
source(paste(dropbox_dir, 'Projects/eb/eb/an/funs_an/BF_t.R', sep='/'))
source(paste(dropbox_dir, 'Projects/eb/eb/an/funs_an/BF_U.R', sep='/'))
## Data input
ds1 = read.csv(paste(dropbox_dir, 'Projects/eb/eb/an/consolidated/ds1.csv', sep='/'))
dsa1 = read.csv(paste(dropbox_dir, 'Projects/eb/eb/an/consolidated/dsa1.csv', sep='/'))
ds2 = read.csv(paste(dropbox_dir, 'Projects/eb/eb/an/consolidated/ds2.csv', sep='/'))
dsa2 = read.csv(paste(dropbox_dir, 'Projects/eb/eb/an/consolidated/dsa2.csv', sep='/'))
```

# Linear models

```{r}
## Only including the predictive cueing condition from Expt 1:
ds1 = ds1[ds1$cue_pred==1,]
## Only including the trials within a certain CTOA range from Expt 2:
ds2 = ds2[ds2$ctoa <= max(ds1$ctoa),]
## Combining the two datasets:
ds = rbind(ds1, ds2)
## A better-named experiment column:
ds$expt = 'Expt1'
ds$expt[ds$exp_name=='eb2'] = 'Expt2'
## Sanity check
## Reordering the levels
ds$Cond <- factor(ds$Cond, c('Prompted','NoBlink','Artificial')) #this works for lm
```

## Sanity check model

A model confirming the effects of Expt 1.

```{r}
ds1$Cond <- factor(ds1$Cond, c('Prompted','NoBlink','Artificial')) #this works for lm
summary(lmer(RT~Cond*cue_valid + (1|subj), data=ds1))
```

## Model with CTOA

```{r}
summary(lmer(RT~expt*Cond*cue_valid*ctoa + (1|subj), data=ds))
```

The model yields only one effect of significance: there is a significant increase in RT in Expt 2. Since there is no interaction with this effect, this applies across eye blink conditions and cue validities.

The fact that there is no effect of CTOA is good, because we can simply ignore it from now on... Although there *is* a marginally significant interaction of Expt# & CTOA, where CTOA in the second experiment is slightly more negative -- implying slightly faster RTs in later time periods.

## Model without CTOA

```{r}
summary(lmer(RT~expt*Cond*cue_valid + (1|subj), data=ds))
```

This model gives the same effect as the previous model, but also:
1) RT was faster on valid trials in Expt 1 than invalid trials in PB condition, confirming the results of Expt 1 (see the sanity check model above),
2) but this effect vanished in the No Blink condition
3) and *reversed* in Expt 2: the slope is now positive (~17 ms), implying IOR across blink conditions.

# Visualization

## Comparing Both Experiments

```{r, fig.width=5.5, fig.height=2.5}
sumss = ddply(ds, .(subj, expt, Cond, CueValid), summarise, RT=median(RT))
p = ggplot(data=sumss, aes(x=Cond, y=RT, fill=CueValid)) + 
    geom_boxplot() + facet_grid(.~expt) + xlab('Eye Blink Condition') + 
    ylab('RT (ms)') + theme_bw() + 
    # geom_dotplot(binaxis='y', stackdir='center', dotsize=2, binwidth=3, ) + 
    theme(panel.grid.minor=element_blank(), 
          plot.title = element_text(hjust = 0.5)) + ggtitle('Experiments 1 & 2 Comparison')
plot(p)
pdf('eb_box_expts1-2.pdf', width=5.5, height=2.5)
plot(p)
dev.off()
```