---
title: "[eb1] Analyses"
author: "Egor Ananyev"
date: "2019-01-04"
output:
  html_notebook: 
    toc: yes
    html_document:
        toc: true
editor_options: 
  chunk_output_type: inline
---

<style type="text/css">
body, td{
  font-family: Cambria;
}
code.r{
  font-family: Inconsolata;
}
</style>

# Data and function input
```{r set-options, message=F}
options(width=100)
library(plyr)
library(dplyr)
library(ggplot2)
library(lme4)
computer = 'acer'  # 'shocky', 'station3', 'acer' are all possible options
if(computer == 'shocky' || computer == 'toshi'){ dropbox_dir='/Users/Egor/Dropbox/' }
if(computer == 'acer'){ dropbox_dir='/Users/egora/Dropbox/' }
source(paste(dropbox_dir, 'Projects/eb/eb/an/funs_an/BF_t.R', sep='/'))
source(paste(dropbox_dir, 'Projects/eb/eb/an/funs_an/BF_U.R', sep='/'))
## Data input
ds = read.csv(paste(dropbox_dir, 'Projects/eb/eb/an/consolidated/ds1.csv', sep='/'))
dsa = read.csv(paste(dropbox_dir, 'Projects/eb/eb/an/consolidated/dsa1.csv', sep='/'))
```

# Cueing

```{r, fig.width=3, fig.height=2.5}
# for(cur_cue in unique(ds$cue_pred)){
#   for(cur_subj in unique(ds$subj)){
#     for(cur_cond in c('a','c','v')){
#       cur_ds = ds[ds$cond==cur_cond & ds$subj==cur_subj & ds$cue_pred==cur_cue,]
#       cur_cond_name = unique(cur_ds$CondFull)
#       p = ggplot(data=cur_ds, aes(x=factor(cue_valid), y=RT))
#       p = p + theme_bw() + ylim(short_rt_cutoff, 900)
#       p = p + geom_dotplot(binaxis='y', stackdir='center', dotsize=1, binwidth=.02)
#       p = p + ggtitle(paste('cond', cur_cond_name))
#       suppressWarnings(plot(p))
#     }
#   }
# }
```

## Individual data
```{r, fig.width=3, fig.height=2.5}
# for(cur_cue in unique(ds$cue_pred)){
#   print('-----------------------------------------------------------------------------------')
#   print(paste('cue predictive?', as.character(cur_cue)))
#   for(cur_subj in unique(ds$subj)){
#     for(cur_cond in c('a','c','v')){
#       cur_ds = ds[ds$cond==cur_cond & ds$subj==cur_subj & ds$cue_pred==cur_cue,]
#       cur_plot_name = paste0(unique(cur_ds$CondFull), ': ', 'Participant ',
#                              as.character(cur_subj))
#       p = ggplot(data=cur_ds, aes(x=RT, colour=factor(cue_valid))) + geom_density() +
#         theme(legend.position = 'bottom') + ggtitle(cur_plot_name)
#       plot(p)
#     }
#   }
# }
```

## Cueing group plots

### Density plots
```{r, fig.width=6, fig.height=2.5}
plot_dens = function(cue_pred_val, plot_title){
  ggplot(data=ds[ds$cue_pred==cue_pred_val,], aes(x=RT, colour=CueValid)) + 
    geom_density() + facet_grid(.~CondFull) + theme(legend.position = 'bottom') + 
    ggtitle(plot_title) + guides(colour=guide_legend(title='Cue Validity'))
}
plot(plot_dens(0, 'Unpredictive Cue'))
plot(plot_dens(1, 'Predictive Cue'))
```

### CTOA

```{r, fig.width=3.5, fig.height=3.5}
ctoa_valid = ddply(ds[ds$CueValid=='Valid',], .(CuePredFull, ctoa_lowbound, Cond, subj), 
                   summarise, RT=median(RT))
plot_box_ctoa(ctoa_valid, 'RT Valid (ms)', 'RT on Valid Trials')
ctoa_invalid = ddply(ds[ds$CueValid=='Invalid',], .(CuePredFull, ctoa_lowbound, Cond, subj), 
                     summarise, RT=median(RT))
plot_box_ctoa(ctoa_invalid, 'RT Invalid (ms)', 'RT on Invalid Trials')
ctoa_adv = merge(ctoa_valid, ctoa_invalid, by=colnames(ctoa_valid)[-5])
ctoa_adv$RT = ctoa_adv$RT.y - ctoa_adv$RT.x  # Invalid - Valid RT
plot_box_ctoa(ctoa_adv, 'RT Invalid - Valid (ms)', 'RT Cueing Advantage')
```

### Cueing advantage (invalid - valid cue)
```{r, fig.width=4.5, fig.height=2.5}
sumss_valid = ddply(ds[ds$CueValid=='Valid',], .(subj, CuePredFull, Cond), summarise, 
                    RT_valid=median(RT))
sumss_invalid = ddply(ds[ds$CueValid=='Invalid',], .(subj, CuePredFull, Cond), summarise, 
                      RT_invalid=median(RT))
sumss_adv = merge(sumss_valid, sumss_invalid, by=c('subj', 'CuePredFull', 'Cond'))
sumss_adv$rt_adv = sumss_adv$RT_invalid - sumss_adv$RT_valid
p = ggplot(data=sumss_adv, aes(x=Cond, y=rt_adv, fill=Cond, alpha=0.5)) + 
    geom_boxplot() + facet_grid(.~CuePredFull) + xlab('Eye Blink Condition') + 
    ylab('RT Invalid - Valid (ms)') + theme_bw() + ylim(-75, 75) +
    geom_dotplot(binaxis='y', stackdir='center', dotsize=2, binwidth=3) + 
    theme(legend.position='none', panel.grid.minor=element_blank(), 
          plot.title = element_text(hjust = 0.5)) + ggtitle('Cueing Advantage')
plot(p)
pdf('eb1_box_adv_med.pdf', width=4.5, height=2.5)
plot(p)
dev.off()
```

## Linear analyses

### Prompted vs. everything
```{r}
# Reordering levels of Condition variable, such that No Blink condition is the reference:
ds$Cond <- factor(ds$Cond, c('Prompted','NoBlink','Artificial')) #this works for lm
summary(lmer(RT~Cond*cue_valid*cue_pred + (1|subj), data=ds))
summary(lm(RT~Cond*cue_valid_c*cue_pred_c, data=ds))
```

### No Blink vs everything
```{r}
ds$Cond <- factor(ds$Cond, c('NoBlink','Artificial','Prompted')) #this works for lm
summary(lmer(RT~Cond*cue_valid*cue_pred + (1|subj), data=ds))
summary(lm(RT~Cond*cue_valid*cue_pred, data=ds))
summary(lm(RT~Cond*cue_valid_c*cue_pred_c, data=ds))
```

## Bayes Factor
```{r}
library(BayesFactor)
bf_ss = ds
bf_ss$subj <- factor(bf_ss$subj)
bf_ss$AB = 0
bf_ss$AB[bf_ss$cond=='a'] = 1
bf_ss$PB = 0
bf_ss$PB[bf_ss$cond=='v'] = 1
bf_ss$NB = 0
bf_ss$NB[bf_ss$cond=='c'] = 1
```

### Predictive cue

#### Prompted vs no blink
```{r}
mRes = 1/generalTestBF(RT ~ PB * cue_valid + subj,
                  data=bf_ss[bf_ss$AB==0 & bf_ss$cue_pred==1,], whichRandom='subj', 
                  neverExclude='subj', whichModels='top')
print(mRes)
```

#### Prompted vs artificial blink
```{r}
mRes = 1/generalTestBF(RT ~ PB * cue_valid + subj,
                  data=bf_ss[bf_ss$NB==0 & bf_ss$cue_pred==1,], whichRandom='subj', 
                  neverExclude='subj', whichModels='top')
print(mRes)
```

#### No blink vs artificial blink
```{r}
mRes = 1/generalTestBF(RT ~ AB * cue_valid + subj,
                  data=bf_ss[bf_ss$PB==0 & bf_ss$cue_pred==1,], whichRandom='subj', 
                  neverExclude='subj', whichModels='top')
print(mRes)
```

# Blink parameters

## Cleaning up the combined dataset
```{r}
cue_valid_name = 'Cue Validity'
dsa$cue_valid_full = 'Invalid'
dsa$cue_valid_full[dsa$cue_valid==1] = 'Valid'
cue_pred_name = 'Cue Type'
dsa$cue_pred_full = 'Unpredictive Cue'
dsa$cue_pred_full[dsa$cue_pred==1] = 'Predictive Cue'
dsa$blink_target_asynchrony = 1000 * (dsa$targ_time - dsa$trial_start - dsa$blank_time_end)
dsa$blink_target_asynchrony_c = dsa$blink_target_asynchrony - mean(dsa$blink_target_asynchrony)
dsa$tot_blank_time = 1000 * (dsa$tot_blank_time)
dsa$tot_blank_time_c = dsa$tot_blank_time - mean(dsa$tot_blank_time)
dsa$blank_time_beg = 1000 * (dsa$blank_time_beg)
dsa$blank_time_beg_c = dsa$blank_time_beg - mean(dsa$blank_time_beg)
head(dsa[,c('blank_time_beg', 'cue_time', 'trial_time_beg')])
dsa$blink_latency = dsa$blank_time_beg - 1000 * (dsa$cue_time - dsa$trial_time_beg)
```

## Blink duration

### PB
```{r}
## Linear model:
summary(lmer(RT~tot_blank_time_c*cue_valid_c*cue_pred_c + (1|subj), 
             data=dsa[dsa$Cond=='Prompted',]))
# summary(lm(RT~tot_blank_time_c*cue_valid_c*cue_pred_c, data=dsa[dsa$Cond=='Prompted',]))
summary(lm(RT~tot_blank_time*cue_valid*cue_pred, data=dsa[dsa$Cond=='Prompted',]))
summary(lm(RT~tot_blank_time*cue_valid*cue_pred, data=dsa[dsa$Cond=='Artificial',]))
```

```{r, fig.width=5.5, fig.height=2.5}
bdur_rt = ddply(dsa[dsa$Cond=='Prompted',], .(subj, cue_pred_full, cue_valid_full), summarise,
                RT=mean(RT), bdur=mean(tot_blank_time), bdur_min=min(tot_blank_time),
                bdur_max=max(tot_blank_time), bdur_sd=sd(tot_blank_time), 
                bdur_supra300=length(tot_blank_time>300))
ggplot(bdur_rt, aes(x=bdur, y=RT)) + geom_point(aes(color=factor(cue_valid_full))) + 
  stat_smooth(method='lm', aes(color=factor(cue_valid_full)), alpha=.2) +
  facet_grid(cols=vars(cue_pred_full)) + xlab('Blink Duration (ms)') + ylab('Reaction Time (ms)')+
  guides(colour=guide_legend(title=cue_valid_name)) + theme_bw() + xlim(75, 225) +
  theme(legend.position=c(.09, .75), legend.title=element_blank())
```

#### Individual slopes
```{r, fig.width=5.5, fig.height=2.5}
p = ggplot(dsa, aes(x=tot_blank_time, y=RT)) + #geom_point(aes(color=factor(subj))) + 
  geom_line(stat='smooth', method='lm', aes(linetype=factor(cue_valid_full), color=factor(subj), 
                                            alpha=.01), se=F) +
  geom_line(stat='smooth', method='lm', size=1, aes(linetype=factor(cue_valid_full))) +
  facet_grid(cols=vars(cue_pred_full), rows=vars(Cond)) + 
  xlab('Blink Duration (ms)') + ylab('Reaction Time (ms)')+
  guides(linetype=guide_legend(title=cue_valid_name), color=F, alpha=F) + theme_bw()
pdf('box_corr_slopes.pdf', width=5.5, height=4.0)
plot(p)
dev.off()
```

#### Average slopes
```{r, fig.width=5.5, fig.height=2.5}
p = ggplot(dsa, aes(x=tot_blank_time, y=RT)) + 
  geom_line(stat='smooth', method='lm', size=1, aes(color=factor(cue_valid_full))) +
  facet_grid(cols=vars(cue_pred_full), rows=vars(Cond)) + 
  xlab('Blink Duration (ms)') + ylab('Reaction Time (ms)')+
  guides(color=guide_legend(title=cue_valid_name), alpha=F) + theme_bw()
plot(p)
```

#### Correlations
```{r, fig.width=5.5, fig.height=2.5}
bdur_rt_corr = ddply(dsa, .(Cond, subj, cue_pred_full, cue_valid_full), summarise,
                corr = cor(tot_blank_time, RT))
p = ggplot(data=bdur_rt_corr, aes(x=cue_valid_full, y=corr, fill=cue_valid_full, alpha=0.5)) + 
    geom_boxplot() + facet_grid(Cond~cue_pred_full) + xlab('Cue Validity') + 
    ylab('Pearson\'s Correlation') + theme_bw() + 
    geom_dotplot(binaxis='y', stackdir='center', dotsize=1, binwidth=.05) + 
    theme(legend.position='none', panel.grid.minor=element_blank(), 
          plot.title = element_text(hjust = 0.5)) + ggtitle('RT x Blink Duration Correlation')
pdf('box_corr_vals.pdf', width=5.5, height=3.5)
plot(p)
dev.off()
```

### PB vs AB
```{r, fig.width=5.5, fig.height=3.5}
bdur_rt = ddply(dsa, .(subj, Cond, cue_pred_full, cue_valid_full), summarise,
                bdur = mean(tot_blank_time), RT = mean(RT))
p = ggplot(bdur_rt, aes(x=bdur, y=RT)) + geom_point(aes(color=factor(cue_valid_full))) + 
  stat_smooth(method='lm', aes(color=factor(cue_valid_full)), alpha=.2) +
  facet_grid(cols=vars(cue_pred_full), rows=vars(Cond)) + xlab('Blink Duration (ms)') + 
  ylab('Reaction Time (ms)')+ guides(colour=guide_legend(title=cue_valid_name)) + theme_bw() + 
  xlim(75, 225) + theme(legend.position=c(.11, .89), legend.title=element_blank(),
                        legend.box.background = element_rect(colour = "black"),
                        legend.spacing.y = unit(0, "mm"))
## Linear model:
# summary(lm(RT~tot_blank_time*cue_valid*cue_pred, data=dsa))
# summary(lmer(RT~tot_blank_time*cue_valid*cue_pred*Cond + (1|subj), data=dsa))
pdf('bdur_pb-ab.pdf', width=4.7, height=3.7)
plot(p)
dev.off()
summary(lmer(RT~tot_blank_time_c*cue_valid_c*cue_pred_c*Cond + (1|subj), data=dsa))
```

## Blink latency

### PB
```{r, fig.width=5.5, fig.height=2.5}
blat_rt = ddply(dsa[dsa$Cond=='Prompted',], .(subj, cue_pred_full, cue_valid_full), summarise,
                blat = mean(blank_time_beg), RT = mean(RT))
p = ggplot(blat_rt, aes(x=blat, y=RT)) + geom_point(aes(color=factor(cue_valid_full))) + 
  stat_smooth(method='lm', aes(color=factor(cue_valid_full)), alpha=.2) +
  facet_grid(cols=vars(cue_pred_full)) + xlab('Blink Latency (ms)') + ylab('Reaction Time (ms)') +
  guides(colour=guide_legend(title=cue_valid_name)) + theme_bw() + 
  theme(legend.position=c(.09, .75), legend.title=element_blank(), 
        legend.box.background = element_rect(colour = "black"),
        legend.spacing.y = unit(0, "mm"))
pdf('blat_pb.pdf', width=5.5, height=2.2)
plot(p)
dev.off()
## Linear model:
summary(lmer(RT~blank_time_beg_c*cue_valid_c*cue_pred_c + (1|subj), 
             data=dsa[dsa$Cond=='Prompted',]))
summary(lm(RT~blank_time_beg_c*cue_valid_c*cue_pred_c, data=dsa[dsa$Cond=='Prompted',]))
```

### AB
```{r, fig.width=5.5, fig.height=2.5}
blat_rt = ddply(dsa[dsa$Cond=='Artificial',], .(subj, cue_pred_full, cue_valid_full), summarise,
                blat = mean(blank_time_beg), RT = mean(RT))
p = ggplot(blat_rt, aes(x=blat, y=RT)) + geom_point(aes(color=factor(cue_valid_full))) + 
  stat_smooth(method='lm', aes(color=factor(cue_valid_full)), alpha=.2) +
  facet_grid(cols=vars(cue_pred_full)) + xlab('Blink Latency (ms)') + ylab('Reaction Time (ms)') +
  guides(colour=guide_legend(title=cue_valid_name)) + theme_bw() + 
  theme(legend.position=c(.09, .75), legend.title=element_blank(), 
        legend.box.background = element_rect(colour = "black"),
        legend.spacing.y = unit(0, "mm"))
pdf('blat_ab.pdf', width=5.5, height=2.2)
plot(p)
dev.off()
## Linear model:
summary(lmer(RT~blank_time_beg_c*cue_valid_c*cue_pred_c + (1|subj), 
             data=dsa[dsa$Cond=='Artificial',]))
summary(lm(RT~blank_time_beg_c*cue_valid_c*cue_pred_c, data=dsa[dsa$Cond=='Artificial',]))
```

## Blink-target asynchrony

### PB
```{r, fig.width=5.5, fig.height=2.5}
bota_rt = ddply(dsa[dsa$Cond=='Prompted',], .(subj, cue_pred_full, cue_valid_full), summarise,
                bota = mean(blink_target_asynchrony), RT = mean(RT))
p = ggplot(bota_rt, aes(x=bota, y=RT)) + geom_point(aes(color=factor(cue_valid_full))) + 
  stat_smooth(method='lm', aes(color=factor(cue_valid_full)), alpha=.2) +
  facet_grid(cols=vars(cue_pred_full)) + xlab('Blink Offset to Target Asynchrony (ms)') + 
  ylab('Reaction Time (ms)') + guides(colour=guide_legend(title=cue_valid_name)) + theme_bw() +
  theme(legend.position='none')
pdf('bota_pb.pdf', width=5.5, height=2.2)
plot(p)
dev.off()
## Linear model:
summary(lmer(RT~blink_target_asynchrony*cue_valid_c*cue_pred_c + (1|subj), 
             data=dsa[dsa$Cond=='Prompted',]))
summary(lm(RT~blink_target_asynchrony*cue_valid_c*cue_pred_c, data=dsa[dsa$Cond=='Prompted',]))
```

### AB
```{r, fig.width=5.5, fig.height=2.5}
bota_rt = ddply(dsa, .(subj, Cond, cue_pred_full, cue_valid_full), summarise,
                bota = mean(blink_target_asynchrony), RT = mean(RT))
ggplot(bota_rt, aes(x=bota, y=RT)) + geom_point(aes(color=factor(cue_valid_full))) + 
  stat_smooth(method='lm', aes(color=factor(cue_valid_full)), alpha=.2) +
  facet_grid(cols=vars(cue_pred_full), rows=vars(Cond)) + 
  xlab('Blink Offset to Target Asynchrony (ms)') + 
  ylab('Reaction Time (ms)') + guides(colour=guide_legend(title=cue_valid_name)) + theme_bw() +
  theme(legend.position='none')
## Linear model:
summary(lmer(RT~blink_target_asynchrony*cue_valid_c*cue_pred_c + (1|subj), 
             data=dsa[dsa$Cond=='Artificial',]))
summary(lm(RT~blink_target_asynchrony*cue_valid_c*cue_pred_c, data=dsa[dsa$Cond=='Artificial',]))
```

## Blink summary stats
```{r}
(ddply(dsa[dsa$Cond=='Prompted',], .(subj), summarise,
      bdur=mean(tot_blank_time), bdur_sd=sd(tot_blank_time),
      bdur_min=min(tot_blank_time), bdur_max=max(tot_blank_time), 
      bdur_sub300=sum(tot_blank_time<300), bdur_supra300=sum(tot_blank_time>300),
      total=length(tot_blank_time)))
(ddply(dsa[dsa$Cond=='Prompted',], .(subj), summarise,
      blat=mean(blink_latency), blat_sd=sd(blink_latency),
      blat_min=min(blink_latency), blat_max=max(blink_latency), 
      blat_sub300=sum(blink_latency<300), blat_supra300=sum(blink_latency>300),
      total=length(blink_latency)))
```