---
title: '[eb4] Analyses'
author: "Egor Ananyev"
date: '2020-10-22'
output:
  html_document: 
    df_print: kable
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: yes
editor_options:
  chunk_output_type: inline
---

<style type="text/css">
body, td{
  font-family: Cambria;
}
code.r{
  font-family: Inconsolata;
}
</style>

# Data input and parsing
```{r set-options, message=F}
options(width=100)
out=F
library(plyr)
library(dplyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(reshape)
library(R.utils)
computer = 'toshi'  # 'shocky', 'station3', 'acer' are all possible options
if(computer=='acer'){
  cloud_dir = 'C:/gd/'
  prog_dir = 'C:/Users/egora/Prog/'
}
if(computer=='shocky'){
  cloud_dir = 'C:/gd/'
  prog_dir = 'C:/Users/Egor/Prog/'
}
if(computer=='station3'){
  cloud_dir = '/home/station3/Desktop/Egor/'
  prog_dir = cloud_dir
}
if(computer == 'toshi') {
  cloud_dir = 'C:/Users/Egor/Google Drive/'
  prog_dir = 'C:/Users/Egor/Prog/'
}
sourceDirectory(paste0(prog_dir, 'eb/an/funs_an'))
source(paste0(cloud_dir, 'Prog/R/myFunctions/pvalfn.R'))
ds = read.csv(paste0(prog_dir, 'eb/an/consolidated/ds4.csv'))
dsa = read.csv(paste0(prog_dir, 'eb/an/consolidated/dsa4.csv'))
# Reordering levels of Condition variable, such that No Blink condition is the reference:
ds$Cond = factor(ds$Cond, c('NoBlink','Artificial','Prompted'))
ds$CondFull = factor(ds$CondFull, c('No Blink','Artificial Blink','Prompted Blink'))
cue_valid_name = 'Cue Validity'
```


# Cueing


## Boxplots
```{r, fig.width=8.5, fig.height=2.5, results="hide"}
ss_corr = ds[ds$corr_resp==1, ]
ctoa_valid = ddply(ss_corr[ss_corr$CueValid=='Valid',], .(targ_soa, Cond, subj), summarise,
                   RT=median(RT))
plot_box_ctoa(ctoa_valid, 'RT Valid (ms)', 'RT on Valid Trials', out, 'targ_soa', eb='eb3')
```

```{r, fig.width=8.5, fig.height=2.5, results="hide"}
ctoa_invalid = ddply(ss_corr[ss_corr$CueValid=='Invalid',], .(targ_soa, Cond, subj), summarise, 
                  RT=median(RT))
plot_box_ctoa(ctoa_invalid, 'RT Invalid (ms)', 'RT on Invalid Trials', out, 'targ_soa', eb='eb3')
```

```{r, fig.width=8.5, fig.height=2.5, results="hide"}
ctoa_adv = merge(ctoa_valid, ctoa_invalid, by=c('targ_soa', 'Cond', 'subj'))
ctoa_adv$RT = ctoa_adv$RT.y - ctoa_adv$RT.x  # Invalid - Valid RT
plot_box_ctoa(ctoa_adv, 'RT Invalid - Valid (ms)', 'RT Cueing Advantage', out, 'targ_soa', eb='eb3')
```

## Analyses

It makes no sense to center cue validity.

A potential rationale for centering validity is to look at the rest of the effects with respect to a "neutral" cue. But this goes against the usefulness of centering: it is meant to focus the effects on *meaningful* levels of a variable, whereas no neutral cue exists in the data. A more logical "reference" level for the valid cue is the invalid one.

For the CTOA, instead of centering on the mean, I went for centering on the shortest asynchrony, and scaling the data such that a unit of 1 is the step from the shortest to longest asynchrony. This way, an effect of CTOA would indicate the change in RT when the asynchrony changes from shortest to longest.

```{r}
ds_rt = ss_corr
ds_rt$ctoa_0 = (ds_rt$targ_soa - 0.3) * 5  # [0.3, 0.4, 0.5] -> [0, 0.1, 0.2] -> [0, 0.5, 1]
# ds_rt$ctoa_c = (ds_rt$targ_soa - 0.4) * 10  # to turn into [-1, 0, 1]
# ds_rt$cue_valid_c = ds_ans$cue_valid * 2 - 1  # to turn [0, 1] into [-1, 1]
```

```{r}
print(lmerfn(lmer(RT ~ cue_valid * Cond * ctoa_0 + block + (1|subj), data = ds_rt)))
# print(anova(lmer(RT ~ cue_valid * Cond * ctoa_0 + (1|subj), data = ds_rt)))
```

The analysis reveals that the RTs for the AB condition are slower than in the no-blink condition (#2). These are only somewhat slower for the prompted condition (#3), with the effect being barely significant.

However, there is no significant difference of cue validity: having a valid cue does speed up the reaction times (#1), but this is unaffected by the blink condition.

```{r}
ds_rt_pb = ds_rt 
ds_rt_pb$Cond = factor(ds_rt_pb$Cond, c('Prompted','Artificial','NoBlink'))
print(lmerfn(lmer(RT ~ cue_valid * Cond * ctoa_0 + block + (1|subj), data = ds_rt_pb)))
```

# Eye blinks

There are several events taking place: blink cue (cue_delay), spatial cue (scue_delay), blink onset and offset (leading to blink duration), target appearance (CTOA), and subject response (RT). These also yield the blink onset relative to the blink cue (blink latency) and relative to the spatial cue (CBOA = cue-blink onset asynchrony, which may be negative).

Longer RTs are associated with longer blinks, but only for later targets. This may hint at the effects of fatigue.

Longer RTs are also associated with /later/ blinks, although the effect is reduced for later targets. Blink latency may be indicative of the individual differences in reaction times. After all, RTs should be similar to targets and blink cues.

Neither effect seems to have anything to do with spatial cue validity: although all analyses pick up the cue validity effect, none of the blink conditions seem to interact with it. That means that blinking does not seem to affect spatial attentional allocation based on an audio spatial cue.


## Blink durations

```{r}
dsa$bdur = dsa$tot_blank_time
bdur_sumss = ddply(dsa, .(Cond, subj), summarize, 
                        bdur_ave = mean(bdur),
                        bdur_sd = sd(bdur))
dsa = merge(dsa, bdur_sumss, by=c('Cond', 'subj'))
dsa$bdur_z = with(dsa, (bdur - bdur_ave) / bdur_sd)
dsa$bdur_c = with(dsa, (bdur - mean(bdur)))
dsa$ctoa_0 = (dsa$targ_soa - 0.3) * 5  # [0.3, 0.4, 0.5] -> [0, 0.1, 0.2] -> [0, 0.5, 1]
```

```{r, fig.width=5.5, fig.height=2.5}
p = ggplot(dsa, aes(x=bdur, y=RT, color=CueValid)) + 
    stat_smooth(method='lm', alpha=0.15) +
    facet_grid(Cond~targ_soa) + guides(color=guide_legend(title='Cue Validity')) +
    xlab('Blink Duration (s)') + ylab('Reaction Time (ms)')
p = plot_themefy(p)
plot(p)
```

```{r}
# print(lmerfn(lmer(RT ~ cue_valid * Cond * ctoa_0 * bdur + block + (1|subj), data = dsa)))
```

Note: Important to center blink duration: otherwise, the model picks up the effects of asynchronies (#3,8) at meaningless blink duration of zero. Trying with a centered, non-normalized blink duration:

```{r}
print(lmerfn(lmer(RT ~ cue_valid * Cond * ctoa_0 * bdur_c + block + (1|subj), data = dsa)))
```

This model picks up the cue validity effect (#1), but this effect does not interact with the blink condition effects.

Another (barely significant) effect is the interaction between the asynchrony and blink durations: longer blink durations increase RT, but only to later targets (#11). In other words, for later targets, an increase of 100 ms in blink durations leads to ~16.7 ms slowdown in RT.


## Blink latencies

```{r}
dsa$blat = dsa$blank_time_beg - (dsa$cue_time - dsa$trial_time_beg)
blat_sumss = ddply(dsa, .(Cond, subj), summarize, 
                        blat_ave = mean(blat),
                        blat_sd = sd(blat))
dsa = merge(dsa, blat_sumss, by=c('Cond', 'subj'))
dsa$blat_c = with(dsa, (blat - mean(blat)))
dsa$blat_z = with(dsa, (blat - blat_ave) / blat_sd)
```

```{r, fig.width=5.5, fig.height=2.5}
p = ggplot(dsa, aes(x=blat, y=RT, color=CueValid)) + 
    stat_smooth(method='lm', alpha=0.15) +
    facet_grid(Cond~targ_soa) + guides(color=guide_legend(title='Cue Validity')) +
    xlab('Blink Latency (s)') + ylab('Reaction Time (ms)')
p = plot_themefy(p)
plot(p)
```

```{r}
lmerfn(lmer(RT ~ cue_valid * Cond * ctoa_0 * blat_c + block + (1|subj), data = dsa))
```

This model picks up the cue validity effect (#1).

Oddly, it also says that, when both artificial and real blink latencies are centered, the real blinks are associated with slightly longer reaction times. This suggests that, in the the simple model, the mean of blink latencies is not the same as here. Might have something to do with the fact that, here, the no-blink condition is not analyzed.

Unlike with blink durations, blink latencies are generally associated with a robust effect on RT: later blinks are associated with much slower reaction times (#4). For later targets, this effect is significantly reduced (#11).
